\section{Introduction}

\subsection{Motivation and Research Question}

Accurately forecasting financial market volatility is essential for risk management, derivatives pricing, and portfolio allocation. Since \citet{engle1982autoregressive} and \citet{bollerslev1986generalized} introduced ARCH and GARCH models, these parametric approaches have dominated empirical volatility forecasting due to their parsimony and ability to capture volatility clustering—the empirical regularity that large returns tend to be followed by further large returns. Nevertheless, GARCH models suffer from a fundamental limitation: they are purely backward-looking, extracting information only from historical price dynamics. They ignore real-time informational shocks transmitted through news media.

Recent advances in natural language processing (NLP) have made it feasible to extract market-relevant sentiment from news in real time. This raises an empirical question: does financial news sentiment contain incremental information for predicting future volatility beyond what GARCH's price history alone can capture? For emerging markets like Vietnam, this question is particularly salient. The Ho Chi Minh City Stock Exchange (HoSE) is dominated by retail investors—individuals with less sophisticated risk models and higher sensitivity to media narratives. Yet Vietnamese financial news exhibits pronounced positive linguistic bias, with growth-oriented framing common even during market downturns. This market structure and linguistic environment suggest that sentiment-augmented models may improve forecasts more substantially in Vietnam than in developed markets.

This paper investigates whether financial news sentiment, extracted from Vietnamese financial news using recent NLP techniques, improves volatility forecasts of the VN-Index (the HoSE benchmark) beyond a standard GARCH(1,1) baseline. Specifically, we ask: \textit{Does daily Vietnamese financial news sentiment, extracted using PhoBERT from CafeF articles, improve next-day realized volatility forecasts of the VN-Index beyond a pure GARCH(1,1) baseline when incorporated through a two-stage GARCH+LSTM hybrid framework?}

\subsection{Limitations of Linear Sentiment-GARCH Models}

A straightforward approach to incorporating sentiment is to augment the GARCH variance equation with contemporaneous or lagged sentiment (a ``GARCH-X'' specification). However, recent evidence suggests that the sentiment-variance relationship is inherently nonlinear. Léber and Egyed (2025) demonstrate using nonlinear Granger causality tests that the relationship between stock market variance and sentiment indices is predominantly nonlinear, not linear. This nonlinearity justifies why sentiment-augmented GARCH-X models often fail to improve forecasts: linear sentiment terms cannot capture state-dependent or threshold effects. 

Deep learning models, particularly Long Short-Term Memory (LSTM) networks, are well-suited to capture such nonlinear temporal dynamics. The key innovation of this paper is to combine GARCH's linear, parsimonious baseline with an LSTM that learns residual corrections in a two-stage framework. Stage 1 fits a baseline GARCH(1,1) to capture linear conditional heteroskedasticity and volatility clustering. Stage 2 trains an LSTM on GARCH residuals, which may contain remaining nonlinear structure, using lagged returns, volatility components, and daily sentiment aggregates as inputs. This hybrid architecture allows the GARCH component to preserve interpretability and efficiency, while the LSTM learns flexible nonlinear sentiment interactions.

\subsection{Data and Main Contribution}

This paper combines a ten-year panel of VN-Index prices (2015--2024) with a corpus of 86,179 financial articles from CafeF.vn. Each article is classified as positive, negative, or neutral using PhoBERT, a Vietnamese language model specifically pre-trained on financial texts. Daily sentiment aggregates and news volume are constructed using a trading-day alignment protocol. The result is a novel dataset linking Vietnamese financial news to index-level volatility.

Our contributions are fourfold:
\begin{enumerate}
    \item We build a Vietnamese financial news sentiment pipeline using PhoBERT applied to a large corpus of CafeF articles, creating the first publicly analyzable link between Vietnamese financial news and VN-Index volatility.
    \item We propose and implement a two-stage GARCH-LSTM hybrid architecture, combining parametric econometric modeling (Stage 1) with deep learning residual correction (Stage 2), rather than replacing GARCH with a fully neural approach.
    \item We empirically validate that the hybrid model improves volatility forecasts relative to a GARCH(1,1) baseline and that simpler linear GARCH-X alternatives perform worse, supporting the hypothesis that sentiment-variance nonlinearity is the mechanism driving improvement.
    \item We investigate whether negative sentiment asymmetrically drives volatility, testing the hypothesis from \citet{vu2023sentiments} that negative news affects volatility in Vietnam while positive news does not.
\end{enumerate}

\subsection{Paper Organization}

The remainder of this paper is organized as follows. Section 2 reviews volatility forecasting literature, financial sentiment analysis, and emerging market microstructure, positioning our contributions relative to prior work. Section 3 describes data construction, the PhoBERT sentiment pipeline, realized volatility definitions, mathematical specifications of GARCH(1,1), GARCH-X, and hybrid GARCH-LSTM architectures, and estimation protocols. Section 4 reports empirical results, diagnostics, baseline specifications, forecast performance, regime-dependent analysis, and robustness checks. Section 5 discusses mechanisms and limitations, connects findings to prior work, and proposes future research directions before concluding.
