\section{Review of Literature}

\subsection{Volatility Modeling: From ARCH to Hybrid Approaches}

The empirical study of financial market volatility began with \citet{engle1982autoregressive}, who introduced autoregressive conditional heteroskedasticity (ARCH) models to capture the phenomenon that large price movements tend to cluster in time. Engle's innovation was to parameterize the conditional variance as a function of lagged squared innovations, allowing volatility to be forecastable from recent shocks. \citet{bollerslev1986generalized} generalized this framework to GARCH models, permitting the conditional variance to depend on both past squared residuals and past conditional variances. The parsimony of GARCH(1,1)—requiring only three parameters in the variance equation—combined with its strong empirical performance across asset classes and time periods, made it the workhorse model for volatility forecasting.

However, parametric GARCH models impose a rigid linear structure: conditional variance is a deterministic, linear function of past returns and variances. This linearity has two limitations. First, GARCH models ignore exogenous information flows (e.g., news, macroeconomic shocks) and thus miss informational content that markets price in real time. Second, the relationship between news information and conditional variance may be nonlinear: the effect of news may depend on regime, sentiment direction, or market state in ways that linear parameterizations cannot capture. These limitations motivate two extensions: (1) augmenting GARCH with exogenous variables, and (2) replacing or complementing the linear specification with nonlinear learning algorithms.

\subsection{News Sentiment and Market Volatility}

A foundational empirical result is that news volume and sentiment directly predict market volatility. \citet{antweiler2004all} analyzed over 1.5 million messages posted on Yahoo! Finance and Raging Bull for 45 large-cap stocks, using computational linguistics to measure bullishness and bearishness. They found that message volume and sentiment disagreement (proxied by standard deviation) significantly predict realized volatility, even though the effect on directional returns is statistically significant but economically small. This seminal work establishes that information intensity itself—beyond sentiment direction—carries volatility predictive power, motivating the inclusion of daily article counts ($n\_articles$) and sentiment dispersion ($sentiment\_std$) as standalone volatility features.

Building on this foundation, recent work has demonstrated that news sentiment improves volatility forecasts in modern asset classes. \citet{bodilsen2025exploiting} augmented traditional HAR (Heterogeneous Autoregressive) models with RavenPack news sentiment indices for the S\&P 500 and individual stocks. They found that macroeconomic news sentiment significantly enhances volatility predictions, whereas firm-specific news sentiment contributes modestly. Critically, they show that overnight news counts have substantial one-period-ahead predictive power independent of sentiment polarity. These findings justify constructing category-specific macroeconomic sentiment indices and including trading-day-aligned news volume as a predictor.

\subsection{Natural Language Processing for Financial Sentiment}

Extracting sentiment from financial text requires specialized NLP models that understand domain-specific language. For English, models like FinBERT have been pre-trained on financial corpora and achieve high accuracy on financial sentiment tasks. For Vietnamese, no such model existed until \citet{nguyen2020phobert}, who introduced PhoBERT—a large-scale pre-trained language model specifically designed for Vietnamese. PhoBERT-base was trained on 20GB of Vietnamese Wikipedia and news text using the RoBERTa architecture and masked language modeling. The monolingual design (unlike multilingual models like XLM-R) allows PhoBERT to capture Vietnamese-specific linguistic nuances and syntactic structures.

Operationalizing PhoBERT for financial text, \citet{sontung2021stock} fine-tuned the base model on over 1,000 CafeF article titles manually labeled into positive, negative, and neutral categories (with class distribution skewed toward positive: 565 positive, 248 neutral, 187 negative). Their model achieved 93\% title-only classification accuracy. However, title-only sentiment overlooks important information in article bodies. The challenge is managing the 256-token sequence length limit of PhoBERT: full-body text requires truncation. A pragmatic approach—and the one adopted in this paper—is to use titles and lead paragraphs, which typically summarize article content succinctly.

\citet{vu2023sentiments} extended this work by fine-tuning PhoBERT with CNN layers on a larger dataset of 40,000 CafeF and VnEconomy articles, using 2,738 hand-labeled samples, to achieve 81\% binary classification accuracy. They then studied market reactions to news using event study methodology over $\pm5$ and $\pm30$ day windows around positive and negative news announcement dates, finding that return means do not significantly change but negative news significantly increases return variance, while positive news does not. This \textit{sentiment asymmetry} result is pivotal: it suggests that in the Vietnamese retail-dominated market, bad news is a volatility driver, but good news has minimal volatility effect. This motivates investigating $negative\_share$ and $positive\_share$ as separate volatility predictors.

\subsection{Hybrid GARCH-Deep Learning Models and Nonlinear Dynamics}

The nonlinearity of sentiment-variance relationships is well-established. \citet{leber2025nonlinear} study an S\&P 500-level sentiment-augmented GARCH-LSTM hybrid using Twitter and Bloomberg news sentiment (2019--2024). They conduct nonlinear Granger causality tests and find that nonlinear causality from sentiment to variance is prevalent, whereas linear causality is rare. This finding directly justifies replacing linear GARCH-X models (which add sentiment as a linear exogenous term) with nonlinear learners. Their GARCH-LSTM architecture—GARCH for Stage 1 baseline, LSTM for Stage 2 residual learning—demonstrates superior out-of-sample VaR forecasting compared to GARCH-only benchmarks.

The intuition is as follows: GARCH captures the linear, time-series autocorrelation structure of conditional variance from returns alone. The residuals from GARCH may contain remaining structure that is nonlinear—e.g., interaction effects between sentiment and returns that only manifest in certain regimes, or threshold effects whereby sentiment matters more when volatility is already high. An LSTM, processing sequences of multivariate inputs (past returns, past variances, sentiment), can learn such state-dependent mappings without imposing parametric linearity. Because LSTMs are sequence models, they naturally process rolling windows of features and learn temporal dependencies. This architecture is appropriate for volatility residuals that may exhibit regime switches or context-dependent patterns.

\subsection{Vietnamese Equity Market Structure and Microstructure}

The Ho Chi Minh City Stock Exchange (HoSE) differs from developed markets in ways that affect news sentiment's relevance. Retail investors constitute a substantial share of daily trading volume, and they are known to be more influenced by media narratives, social sentiment, and herding behavior than institutional investors. Retail investors often use sentiment-based signals (e.g., news tone) directly in trading decisions without adjusting for statistical bias or market efficiency. This behavioral amplification suggests that news sentiment may have larger volatility effects in Vietnam than in developed markets where institutional investors average sentiment-based trading.

Additionally, Vietnamese financial media exhibits pronounced positive bias. Articles use growth-oriented language even in neutral or negative market periods, inflating the unconditional average sentiment. This positive skew creates a high signal-to-noise ratio: negative news stands out against the baseline positive tone, potentially amplifying its market impact. This media structure further motivates studying asymmetric sentiment effects and motivates the use of categorical sentiment aggregates (macroeconomic vs. market-specific) rather than a single aggregate sentiment scalar.

\subsection{Research Gap and Contribution}

Existing literature establishes three facts: (1) news sentiment predicts volatility in developed markets (Bodilsen \& Lunde, 2025), (2) sentiment-variance relationships are nonlinear and justify LSTM-based learning (Léber \& Egyed, 2025), and (3) Vietnamese news sentiment affects volatility asymmetrically, with negative sentiment mattering but positive sentiment not (Vu et al., 2023). However, no prior work combines these insights in the Vietnamese context: specifically, no paper implements a hybrid GARCH-LSTM model on VN-Index data using PhoBERT sentiment from a large CafeF corpus, nor do existing studies systematically evaluate whether nonlinear learning (LSTM) outperforms linear sentiment-GARCH (GARCH-X) on Vietnamese data.

This paper fills this gap by implementing the hybrid architecture on Vietnamese data, evaluating it against both pure GARCH and linear GARCH-X benchmarks, and investigating asymmetry and regime-dependent effects. Our contribution is not methodological innovation (the GARCH-LSTM framework exists in \citet{leber2025nonlinear}), but rather empirical application to a new market with novel features: a large Vietnamese news corpus, domain-adapted PhoBERT sentiment, a retail-dominated exchange with demonstrated asymmetric sentiment effects, and two crisis periods (COVID-19 in 2020 and liquidity shock in 2022) that allow regime analysis.
